---
title: "Data_structures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Structures
### Vectors
- Collection of homogeneous data
- Coerces(converts) all the members of the vector to higher order data types
- Order of precision(character > complex > numeric > logical)
```{r}
vec = c(2,1,4,5,24,7,18)
vec
char_vec = c('h','e','l','l','o')
char_vec

## mixed input's to a vector
mix_vec = c(2, 'Hey', TRUE)
mix_vec
```



## Generating a sequence
```{r}
my_seq = 1:22

gap_seq = seq(1, 45, 2)
```


### Playing with vectors
```{r}
vec = 6
# numerical operations on a numerical vetor
vec*3
vec/12

vec = c(4,2,31,12,46)

# numerical operations on a numerical vetor
vec*3
vec/12

# Mathematical operations between vectors
vec1 = c(2,1,4,5)
vec2 = c(8,4,1,9)
vec1 + vec2
vec1 - vec2
vec1*vec2
vec1/vec2
vec1%%vec2
vec1%/%vec2

# functions on a vector
length(vec)
nchar(mix_vec)
sort(vec)


# Indexing a vector
vec[2]

# subsetting vectors
vec[3:5]


# logical operations inside a vector
vec == 18
vec[vec == 18]
vec[!vec == 18]
vec[vec > 7]
vec[vec %in% c(2,1,4)]
```


## Lists
- Can hold any data type or data structure
- Ulike vectorslists don't perform coercion to diiferent data types
```{r}
# creating a list
l = list('car',122200,c(1,23,12,87),c(21,3,22))
l

# namming a list
names(l) = c('item','cost','specs','random')
l
names(l)[3] = 'specifications' 
l
names(l)

# accessing list by name
l[['cost']]
l$cost
l$specifications[3]

# Accessing elements of a list
l[1]    #returns a list of that element
l[[1]]   ## returns element at that position
l[1:3]
l[c(1,4,2)]

# print structure of a list
str(l)

# list can contain any complex data type in R
my_list = list(matrix(c(1:6),nrow = 2,byrow = TRUE),c(2,1,8),list(2,'aa'))
my_list
```


## Arrays
- Multi-dimentional collection of data
- Can't hold di
```{r}
#arrays
# creating an array
# 1D array(by default)
my_array = array(1:12)
class(my_array)
my_array

my_array2 = array(1:12, dim = c(4,3))
my_array2

arr = array(c(1,2,3,"a","b","c"),dim = c(2,3))
arr

three_d_array <- array(1:24,dim = c(4, 3, 2))
three_d_array

# accessing elements of an array
my_array2[2,3]
three_d_array[3,1,2]

# subsetting an array
three_d_array[2:4,1:2,2]
three_d_array[,,1]
three_d_array[c(4,2),c(1,2),2]


# Manipulating an array
my_array2[2,3] = 44
my_array

# properties of an array
dim(three_d_array)
nrow(my_array2)
ncol(my_array2)
nrow(three_d_array)
ncol(three_d_array)
```


## Matrices
- 2-dimentional array
- Only holds homogeneous data
```{r}
# creating a matrix
mat = matrix(c(1:24), nrow = 6, byrow = T)
mat

# acessing elements of a matrix
mat[1,2]
mat[4,3]

# indexing a matrix
mat[2:6,1:4]
mat[,3:4]

# naming rows and columns of a matrix
dimnames(mat) = list(c('a','b','c','d','e','f'),c('x','y','z','p'))
mat

# acessing elements of a matrix by row and column names
mat[c('b','e'),c('x','p')]

# properties of a matrix
dim(mat)
nrow(mat)
ncol(mat)
```

























